An interdisciplinary examination of stress and injury occurrence in athletes

### 2020-07-29 apa_style.tex
progress

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  english,
  man]{apa6}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={An interdisciplinary examination of stress and injury occurrence in athletes},
  pdflang={en-EN},
  pdfkeywords={Sports Injury, Stress, Interdisciplinary, Bayesian Network},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
% Manuscript styling
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

@@

Correspondence concerning this article should be addressed to Harry Fisher, Cyncoed Road, Cardiff, CF23 6XD. E-mail: harryfisher21@gmail.com}
\keywords{Sports Injury, Stress, Interdisciplinary, Bayesian Network}
\DeclareDelayedFloatFlavor{ThreePartTable}{table}
\DeclareDelayedFloatFlavor{lltable}{table}
\DeclareDelayedFloatFlavor*{longtable}{table}
\makeatletter
\renewcommand{\efloat@iwrite}[1]{\immediate\expandafter\protected@write\csname efloat@post#1\endcsname{}}
\makeatother
\usepackage{lineno}

@@

\section{Introduction}

@@

@@

\section{Methods}

\subsection{Participants}

A total of 351 athletes (male: \emph{n} = 231, female: \emph{n} = 120) were initially recruited for the study from a British University and local sports clubs (Table \ref{tab:table1}). Participants had an average age of 22.0 \(\pm\) 7.0 years and represented a range of team (football, rugby, netball, cricket, lacrosse, basketball and field hockey) and individual sports (athletics, tennis, weightlifting, gymnastics, judo, swimming and golf.
Participants self-rated competitive level ranged from recreational to international standard.
They were engaged in training for their respective sports for at least five hours per week.
A total of 126 (49.03\%) participants had sustained an injury in the 12 months prior to the start of the study (male: \emph{n} = 114 {[}49\%{]}, female; \emph{n} = 48 {[}40\%{]}).
All participants were injury free (no modifications to their usual training routine due to a sport related medical problem for a minimum of four weeks) at the start of the study.
Ethical approval was obtained from the University ethics committee prior to the start of the study and all participants provided informed consent.

@@

@@

@@

@@

@@

BNs have recently been used in the sport psychology literature (Olmedilla et al., 2018; Fuster-Parra et al., 2017) and offer several benefits over traditional statistical analysis.
For example, predictions can be made about any variable in the network, rather than there being a distinction between dependent and independent variables in the data, such as in linear regression models that are often used within the sport psychology literature (Olmedilla et al., 2018; Bittencourt et al., 2016).
Furthermore, the structure of a network can be obtained from both empirical data \emph{and} prior knowledge about the area of study; the latter being particularly useful when there are a large number of variables in the network, or only a small number of observations are available in the data (Xiao-xuan et al., 2008).
In such instances, a purely data driven approach to learning the network would be time-consuming due to the large parameter space, and inefficiency at identifying an approximation of the true network structure.
Prior knowledge about dependencies between variables can therefore be included in the network structure, while still allowing a data driven approach for unknown dependencies, to improve the overall computation of the network structure (Heckerman et al., 1995; Xu et al., 2015).
The following sections detail the steps taken in the current study to firstly prepare the data for each network, and then obtain the structure of each network that was used for further inference.

@@

A median split technique was used to discretise the data used in the network into ``Low'' and ``High'' levels.
All variables apart from negative and total life events were approximately normally distributed (based on visual inspection, see supplementary material Sx) and required no further transformation prior to the median split.
For the LESCA questionnaire data, a cumulative total of the current, and previous time points was calculated at each time point to account for the potential continuing effect of the life events experienced by athletes over time.
Given the limited support for a relationship between positive life events and injury (Williams and Andersen, 2007), only negative and total life events were included in the network.
Cumulative negative, and cumulative total life event scores at each time point were first log scaled so distributions were approximately normal, and then binarised using the median at each time point (nlelg and tlelg respectively).
In addition to the log scaled cumulative values, an untransformed NLE score from the first time point was included as an additional variable based on previous literature that indicates this variable should have a strong relationship with injury outcome (Ivarsson et al., 2017).

@@

To obtain the network structure, several steps were taken to ensure that both a theoretically realistic network, and a network that was an appropriate fit to the collected data, was used for inference.
Prior knowledge about the network structure was included by providing a list of arcs that are always \emph{restricted} from being in the network (blacklist), and a list of arcs that are always \emph{included} in the network (whitelist).
Additionally, there are several scoring functions such as Bayesian Information Criteria (BIC) and Bayesian Dirichlet equivalent uniform (BDeu) that can be used to compare network structures with certain nodes and arcs included or excluded (Scutari and Denis, 2014).
To account for the repeated measures design employed in this study and to maximise the use of the data, pairs of complete cases (e.g., participants who completed T1 + T2, and T2 + T3) were used in a two-time Bayesian network (2TBN) structure (Murphy, 2002).
In the 2TBN, variables measured T2 could depend on variables measured at T1 (e.g., T1 \(\rightarrow\) T2) and variables measured at T3 could depend on variables measured at T2 (e.g., T2 \(\rightarrow\) T3).
However, arcs were blacklisted between T2 \(\rightarrow\) T1 and T3 \(\rightarrow\) T2 to preserve the order in which data was collected.
Variables were separated into two groups; ``explanatory'', for variables that did not change during the study (e.g., gender), or ``independent'', for variables that were measured at each time point and could vary during the study.
Independent variable names were suffixed with \_1 for time point T, and \_2 for time point T+1 (e.g., T1\_1 \(\rightarrow\) T2\_2 and T2\_1 \(\rightarrow\) T3\_2).
Formatting the data in this way meant participants who completed T1 and T2, but did not complete T3, could still be included in the analysis.
Table \ref{tab:table2} provides an example of the formatted data and demonstrates that participants 1 and 3 have complete data, and therefore have two rows of data each representing variables from T1 \(\rightarrow\) T2 and T2 \(\rightarrow\) T3, respectively.
Participant 2 did not complete the final data collection at T3 and therefore only has one row of data representing the variables collected at T1 and T2.
In addition to the blacklisted arcs between T2 \(\rightarrow\) T1 and T3 \(\rightarrow\) T2, the direction of arcs was restricted between independent variables and explanatory variables (e.g., independent \(\rightarrow\) explanatory); however, arcs were not restricted between explanatory \(\rightarrow\) independent variables.
Finally, arc direction was restricted between specific nodes within the explanatory variables.
Arcs from clevel \(\rightarrow\) gender, nlebase \(\rightarrow\) gender and nlebase \(\rightarrow\) ind\_team were included in the blacklist, as arcs in these directions did not make logical sense.
All subsequent models used the same blacklist.

@@

Initially, both negative life events and total life events were included in the network structure, however, the network score was improved when only nlelg or tlelg was included (highest BIC value = nleleg, BIC values relative to nlelg; tlelg only = -83.64, tlelg and nlelg = -218.31).
Additionally, despite strong evidence in the literature that both negative (NLE) and total life event (TLE) stress are related to injury occurrence (Williams and Andersen, 2007), network structures learned using the Tabu search algorithm failed to identify a relationship between NLE and injury or TLE and injury in the data.
Given that nlelg provided the highest network score, and there is a stronger relationship between negative life events and injury in the literature, an arc was whitelisted between nlelg\_1 and injured\_1 and nlelg\_2 and injured\_2 in the final network structure. TLE score was not included in the final structure.

The subscales representing the BAS (RR, RI, GDP and I) showed limited connection to other variables in the network.
Therefore, several models were run with each scale individually to find the scale that resulted in the highest BIC value (values are shown relative to the highest value).
RI provided the highest BIC value, compared to RR (-9.07), GDP (-15.06) and I (-16.33).
Including all the variables (RR, RI, GDP and I) resulted in a significantly lower score -894.33) indicating that including all the variables was not beneficial to the model structure and did not offset the cost of the additional parameters. Therefore, only RI was included in the final structure.

Finally, both total score and asymmetry for balance were included in the initial network. However, visual inspection of the network revealed no arcs between bal\_asym\_1 or bal\_asym\_2 and any other node in the network.
Therefore, balance asymmetry was removed from the final network structure.
To summarise, Table \ref{tab:table3} reports the variables that were included in the final network structure.

@@

\caption{\label{tab:table3}Variables included in the final Bayesian network structure.}
\centering
\begin{tabular}[t]{l|>{\raggedright\arraybackslash}p{5.2cm}|c|c}
\hline
Variable & Definition & State 1 & State 2\\
\hline
clevel & Current competitive level & Club\_university\_county & National\_international\\
\hline
gender & Gender of the participant & Female & Male\\
\hline
hours & Number of hours spent training per week & 0-9 (Low) & >9-35 (High)\\
\hline
ind\_team & Participate in an individual or team based sport & Individual & Team\\
\hline
pi & Previous injury - Whether an injury had been sustained in the previous 12 months prior to the study & No Injury & Injury\\
\hline
nlebase & Untransformed NLE at TP 1 & 0-13 (Low) & >13-93 (High)\\
\hline
FFFS & Fight-Flight-Freeze System & 8-16 (Low) & >16-30 (High)\\
\hline
BIS & Behavioural Inhibition System & 17-38 (Low) & >38-68 (High)\\
\hline
RI & Reward Interest & 4-10 (Low) & >10-16 (High)\\
\hline
stiffness & Sum of all stiffness locations & 1543-2330 (Low) & >2330-4518 (High)\\
\hline
HRV & Root mean squared difference of successive RR intervals & 2.03-4 (Low) & >4-5.94 (High)\\
\hline
balance & Total balance score & 5-15 (Low) & >15-46 (High)\\
\hline
nlelg\_1 & Log NLE at TP 1 & 0-2.64 (Low) & >2.64-4.54 (High)\\
\hline
nlelg\_2 & Log NLE at TP 2 & 0-3.04 (Low) & >3.04-5.19 (High)\\
\hline
nlelg\_3 & Log NLE at TP 3 & 0-3.18 (Low) & >3.18-4.79 (High)\\
\hline
\end{tabular}
\end{table}

@@

During the study, 26\% of participants reported at least one injury with an average severity of 11 \(\pm\) 31, days (range = 2 - 365 days).
Both male and female participants reported a greater number of acute compared to chronic injuries (male, acute = 85 {[}69\%{]}, chronic = 39 {[}31\%{]}; female, acute = 38 {[}72\%{]} chronic = 15 {[}28\%{]}), and non-contact injuries were more common than contact injuries (male, non-contact = 83 {[}67\%{]}, contact = 39 {[}31\%{]}; female, non-contact = 35 {[}66\%{]} contact = 18 {[}34\%{]}).
Table \ref{tab:table4} shows the number and percentage of injury types sustained by both male and female participants.

@@

The first network structure obtained from the data is shown in Fig \ref{fig:fig6}.
Several of the explanatory variables showed strong connections with independent variables in the network.
The arc from nlebase \(\rightarrow\) RI\_1 had a strength of 0.84, and the probability of RI\_1 being in the ``High'' state increased from 0.23 to 0.47 when nlebase increased from ``Low'' to ``High''.
The ind\_team node had strong arcs to hours (0.90) and nlebase (0.84).
Individual athletes were more likely to have ``High'' hours per week (0.84) compared to team-based athletes (0.60).
Individual athletes were also more likely to have ``High'' negative life events in the 12 months preceding the start of the study compared to team based athletes (individual athletes = 0.65, team-based athletes = 0.41).
The arcs from gender \(\rightarrow\) stiffness\_1 and gender \(\rightarrow\) stiffness\_2 were 0.76, and 0.65 respectively, with males more likely to have ``High'' stiffness compared to females (males = 0.62, females = 0.43).
The arc from pi \(\rightarrow\) stiffness\_1 was 0.55 with athletes who reported an injury in the preceding 12 months more likely to have ``High'' (0.65) compared to ``Low'' (0.35) stiffness.
The arc from clevel \(\rightarrow\) balance\_1 had a strength of 0.51, with lower level performers more likely to have decreased balance ability (0.48), compared to national level athletes (0.29).
Arcs were also present between independent variables in the network. Strong arcs were present between BIS\_1 \(\rightarrow\) FFFS\_1 (0.98) and BIS\_2 \(\rightarrow\) FFFS\_2 (0.68).
In both instances, ``High'' FFFS was more likely when BIS was ``High'' (0.65 for \_1, 0.61 for \_2) compared to ``Low'' (0.32 for \_1, 0.36 for \_2).
The arc between nlelg \(\rightarrow\) BIS had a strength of 0.62 for nlelg\_1 \(\rightarrow\) BIS\_1, however no arc was present between nlelg\_2 and BIS\_2.
For nlelg\_1 \(\rightarrow\) BIS\_1, ``Low'' negative life events increased the probability of BIS being in the ``High'' state from 0.33 to 0.55.

@@

The Markov blanket for injured\_1, which contained hours spent training per week (hours), negative life events (nlelg\_1), muscle stiffness (stiffness\_1), current competitive level (clevel) and balance (balance\_1), is shown in Fig \ref{fig:fig7}.
The arc between nlelg\_1 and injured\_1 was fixed in the network, so has the maximum strength of 1.

The conditional probability query (CPQ) for injured\_1 in the ``injured'' state for all variables that were directly linked to injured\_1 is shown in Table \ref{tab:table5}.
The probability of injured\_1 = ``injured'' rose from 0.17 to 0.31 when stiffness was ``High'' compared to ``Low''.
Negative life events had a negligible effect when moving from the ``Low'' to ``High'' state.

@@

(not sure where this would go /if needed at all?). The first BN revealed both high stiffness and poor balance resulted in an increased probability of injury. However, contrary to previous literature, negative life event stress had a negligible effect on the probability of injury.
In the second network that modelled changes between time points in the study, increases in muscle stiffness was found to increase the risk of injury. In addition, the probability of injury was further increased when there were increases in NLE stress.

(Main discussion starts here..)

>>>>

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  english,
  man,floatsintext]{apa6}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={An interdisciplinary examination of stress and injury occurrence in athletes},
  pdflang={en-EN},
  pdfkeywords={Sports Injury, Stress, Interdisciplinary, Bayesian Network},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
% Manuscript styling
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

@@

Correspondence concerning this article should be addressed to Harry Fisher, Cyncoed Road, Cardiff, CF23 6XD. E-mail: harryfisher21@gmail.com}
\keywords{Sports Injury, Stress, Interdisciplinary, Bayesian Network}
\usepackage{lineno}

@@

10422 words in text body
3142 words in reference section

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

@@

\begin{figure}[!h]

{\centering \includegraphics[width=5.67in]{figures_doc/Fig1}

}

\caption{Stress and injury model (Williams and Andersen, 1998).}\label{fig:fig1}
\end{figure}

@@

\hypertarget{material-and-methods}{%
\section{Material and Methods}\label{material-and-methods}}

\hypertarget{ethics-statement}{%
\subsection{Ethics Statement}\label{ethics-statement}}

Ethical approval was obtained from the University ethics committee prior to the start of the study and all participants provided informed consent.

\hypertarget{participants}{%
\subsection{Participants}\label{participants}}

A total of 351 athletes (male: \emph{n} = 231, female: \emph{n} = 120) were initially recruited for the study from a British University and local sports clubs (Table \ref{tab:table1}). Participants had an average age of 22.0 \(\pm\) 7.0 years and represented a range of team (football, rugby, netball, cricket, lacrosse, basketball and field hockey) and individual sports (athletics, tennis, weightlifting, gymnastics, judo, swimming and golf.
Participants self-rated competitive level ranged from recreational to international standard.
They were engaged in training for their respective sports for at least five hours per week.
A total of 126 (49.03\%) participants had sustained an injury in the 12 months prior to the start of the study (male: \emph{n} = 114 {[}49\%{]}, female; \emph{n} = 48 {[}40\%{]}).
All participants were injury free (no modifications to their usual training routine due to a sport related medical problem for a minimum of four weeks) at the start of the study.

@@

\begin{figure}[!h]

{\centering \includegraphics[width=1\linewidth]{figures_doc/Fig2}

}

\caption{mBESS positions (A-F). Top row, firm surface. Bottom row, foam surface. Left column, dominant leg stance. Middle column, non-dominant leg stance. Right column, Tandem leg stance.}\label{fig:fig2}
\end{figure}

@@

\begin{figure}
\includegraphics[width=1\linewidth]{figures_doc/studydesign_new} \caption{Study design. For each time point (T), each box contains the number of participants who completed the data collection (\textit{n}), the measures used for data collection and the approximate date of the data collection.}\label{fig:fig3}
\end{figure}

@@

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{figures_doc/sessionprotocol}

}

\caption{Outline of the protocol for each data collection.}\label{fig:fig4}
\end{figure}

@@

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{figures_doc/examplenetwork}

}

\caption{A simple discrete Bayesian network contain nodes, possible states of the nodes and the arcs connecting nodes.}\label{fig:fig5}
\end{figure}

@@

BNs have recently been used in the sport psychology literature (Olmedilla et al., 2018; Fuster-Parra et al., 2017; Ruiz-Pérez et al., 2019) and offer several benefits over traditional statistical analysis.
For example, predictions can be made about any variable in the network, rather than there being a distinction between dependent and independent variables in the data, such as in linear regression models that are often used within the sport psychology literature (Olmedilla et al., 2018; Bittencourt et al., 2016).
Furthermore, the structure of a network can be obtained from both empirical data \emph{and} prior knowledge about the area of study; the latter being particularly useful when there are a large number of variables in the network, or only a small number of observations are available in the data (Xiao-xuan et al., 2008).
In such instances, a purely data driven approach to learning the network would be time-consuming due to the large parameter space, and inefficiency at identifying an approximation of the true network structure.
Prior knowledge about dependencies between variables can therefore be included in the network structure, while still allowing a data driven approach for unknown dependencies, to improve the overall computation of the network structure (Heckerman et al., 1995; Xu et al., 2015).
The following sections detail the steps taken in the current study to firstly prepare the data for each network, and then obtain the structure of each network that was used for further inference.

@@

A median split technique was used to discretise the data used in the network into ``Low'' and ``High'' levels.
All variables apart from negative and total life events were approximately normally distributed and required no further transformation prior to the median split.
For the LESCA questionnaire data, a cumulative total of the current, and previous time points was calculated at each time point to account for the potential continuing effect of the life events experienced by athletes over time.
Given the limited support for a relationship between positive life events and injury (Williams and Andersen, 2007), only negative and total life events were included in the network.
Cumulative negative, and cumulative total life event scores at each time point were first log scaled so distributions were approximately normal, and then binarised using the median at each time point (nlelg and tlelg respectively).
In addition to the log scaled cumulative values, an untransformed NLE score from the first time point was included as an additional variable based on previous literature that indicates this variable should have a strong relationship with injury outcome (Ivarsson et al., 2017).

@@

To obtain the network structure, several steps were taken to ensure that both a theoretically realistic network, and a network that was an appropriate fit to the collected data, was used for inference.
Prior knowledge about the network structure was included by providing a list of arcs that are always \emph{restricted} from being in the network (blacklist), and a list of arcs that are always \emph{included} in the network (whitelist).
Additionally, there are several scoring functions such as Bayesian Information Criteria (BIC) and Bayesian Dirichlet equivalent uniform (BDeu) that can be used to compare network structures with certain nodes and arcs included or excluded (Scutari and Denis, 2014).
To account for the repeated measures design employed in this study and to maximise the use of the data, pairs of complete cases (e.g., participants who completed T1 + T2, and T2 + T3) were used in a two-time Bayesian network (2TBN) structure (Murphy, 2002).
In the 2TBN, variables measured at T2 could depend on variables measured at T1 (e.g., T1 \(\rightarrow\) T2) and variables measured at T3 could depend on variables measured at T2 (e.g., T2 \(\rightarrow\) T3).
However, arcs were blacklisted between T2 \(\rightarrow\) T1 and T3 \(\rightarrow\) T2 to preserve the order in which data was collected.
Variables were separated into two groups; ``explanatory'', for variables that were fixed (e.g., gender), or ``independent'', for variables that were measured at each time point and could vary during the study.
Independent variable names were suffixed with \_1 for time point T, and \_2 for time point T+1 (e.g., T1\_1 \(\rightarrow\) T2\_2 and T2\_1 \(\rightarrow\) T3\_2).
Formatting the data in this way meant participants who completed T1 and T2, but did not complete T3, could still be included in the analysis.
Table \ref{tab:table2} provides an example of the formatted data and demonstrates that participants 1 and 3 have complete data, and therefore have two rows of data each representing variables from T1 \(\rightarrow\) T2 and T2 \(\rightarrow\) T3, respectively.
Participant 2 did not complete the final data collection at T3 and therefore only has one row of data representing the variables collected at T1 and T2.
In addition to the blacklisted arcs between T2 \(\rightarrow\) T1 and T3 \(\rightarrow\) T2, the direction of arcs was restricted between independent variables and explanatory variables (e.g., independent \(\rightarrow\) explanatory); however, arcs were not restricted between explanatory \(\rightarrow\) independent variables.
Finally, arc direction was restricted between specific nodes within the explanatory variables.
Arcs from clevel \(\rightarrow\) gender, nlebase \(\rightarrow\) gender and nlebase \(\rightarrow\) ind\_team were included in the blacklist, as arcs in these directions did not make logical sense.
All subsequent models used the same blacklist.

@@

Initially, both negative life events and total life events were included in the network structure, however, the network score was improved when only nlelg or tlelg was included (highest BIC value = nleleg, BIC values relative to nlelg; tlelg only = -83.64, tlelg and nlelg = -216.47).
Additionally, despite strong evidence in the literature that both negative (NLE) and total life event (TLE) stress are related to injury occurrence (Williams and Andersen, 2007), network structures learned using the Tabu search algorithm failed to identify a relationship between NLE and injury or TLE and injury in the data.
Given that nlelg provided the highest network score, and there is a stronger relationship between negative life events and injury in the literature, an arc was whitelisted between nlelg\_1 and injured\_1 and nlelg\_2 and injured\_2 in the final network structure. TLE score was not included in the final structure.

The subscales representing the BAS (RR, RI, GDP and I) showed limited connection to other variables in the network.
Therefore, competing models were examined and BIC scores compared to establish the model with the best fit to the data (values are shown relative to the highest value).
RI provided the highest BIC value (-3,563.13), compared to RR (-15.98), GDP (-19.26) and I (-19.76).
Including all the variables (RR, RI, GDP and I) resulted in a significantly lower score -900.12) indicating that including all the variables was not beneficial to the model structure and did not offset the cost of the additional parameters. Therefore, only RI was included in the final structure.
Finally, both total score and asymmetry for balance were included in the initial network. However, visual inspection of the network revealed no arcs between bal\_asym\_1 or bal\_asym\_2 and any other node in the network.
Therefore, balance asymmetry was removed from the final network structure.
To summarise, Table \ref{tab:table3} reports the variables that were included in the final network structure.

@@

\caption{\label{tab:table3}Variables included in the final Bayesian network structure.}
\centering
\begin{tabular}[t]{l|>{\raggedright\arraybackslash}p{5.2cm}|c|c}
\hline
Variable & Definition & State 1 & State 2\\
\hline
clevel & Current competitive level & Club\_university\_county & National\_international\\
\hline
gender & Gender of the participant & Female & Male\\
\hline
hours & Number of hours spent training per week & 0-9 (Low) & >9-35 (High)\\
\hline
ind\_team & Participate in an individual or team based sport & Individual & Team\\
\hline
pi & Previous injury - Whether an injury had been sustained in the previous 12 months prior to the study & No Injury & Injury\\
\hline
nlebase & Untransformed NLE at TP 1 & 0-13 (Low) & >13-93 (High)\\
\hline
FFFS & Fight-Flight-Freeze System & 8-16 (Low) & >16-30 (High)\\
\hline
BIS & Behavioural Inhibition System & 17-38 (Low) & >38-68 (High)\\
\hline
RI & Reward Interest & 4-10 (Low) & >10-16 (High)\\
\hline
stiffness & Sum of all stiffness locations & 1543-2330 (Low) & >2330-4518 (High)\\
\hline
HRV & Root mean squared difference of successive RR intervals & 2.03-4.01 (Low) & >4.01-5.94 (High)\\
\hline
balance & Total balance score & 5-15 (Low) & >15-46 (High)\\
\hline
nlelg\_1 & Log NLE at TP 1 & 0-2.64 (Low) & >2.64-4.54 (High)\\
\hline
nlelg\_2 & Log NLE at TP 2 & 0-3.04 (Low) & >3.04-5.19 (High)\\
\hline
nlelg\_3 & Log NLE at TP 3 & 0-3.18 (Low) & >3.18-4.79 (High)\\
\hline
\end{tabular}
\end{table}

@@

During the study, 46\% (\emph{n} = 117) of participants reported at least one injury with an average severity of 11 \(\pm\) 31, days (range = 2 - 365 days).
Both male and female participants reported a greater number of acute compared to chronic injuries (male, acute = 85 {[}69\%{]}, chronic = 39 {[}31\%{]}; female, acute = 38 {[}72\%{]} chronic = 15 {[}28\%{]}), and non-contact injuries were more common than contact injuries (male, non-contact = 83 {[}67\%{]}, contact = 39 {[}31\%{]}; female, non-contact = 35 {[}66\%{]} contact = 18 {[}34\%{]}).
Table \ref{tab:table4} shows the number and percentage of injury types sustained by both male and female participants.

@@

The first network structure obtained from the data (Fig \ref{fig:fig6}) examined the interactions between explanatory variables, independent variables and probability of injury across time points in the study.
Strong sport-related and gender-based connections between several explanatory and independent variables were demonstrated for individual and team-based sports.
The ``ind\_team'' node had strong arcs to hours (0.94) and nlebase (0.78).
Individual athletes were more likely to have ``High'' hours per week (0.84) compared to team-based athletes (0.60).
Individual athletes were also more likely to have ``High'' negative life events in the 12 months preceding the start of the study compared to team-based athletes (individual athletes = 0.65, team-based athletes = 0.41).
The arc from clevel \(\rightarrow\) balance\_1 had a strength of 0.47, with lower level performers more likely to have decreased balance ability (0.48), compared to national level athletes (0.29).
High gender-based connections were reported for the arcs from gender \(\rightarrow\) stiffness\_1 (0.71) and gender \(\rightarrow\) stiffness\_2 (0.43), with males more likely to have ``High'' stiffness compared to females (males = 0.62, females = 0.43).
Irrespective of sport or gender, strong connections were found between explanatory variables. The arc from nlebase \(\rightarrow\) RI\_1 had a strength of 0.84, and the probability of RI\_1 being in the ``High'' state increased from 0.47 to 0.77 when nlebase increased from ``Low'' to ``High''.

The first network demonstrated further strong variable interactions between high stiffness, poor balance and injury probability.
The arc from pi \(\rightarrow\) stiffness\_1 was 0.57 with athletes who reported an injury in the preceding 12 months being more likely to have ``High'' (0.65) compared to ``Low'' (0.35). stiffness.
Strong arcs were present between BIS\_1 \(\rightarrow\) FFFS\_1 (0.98) and BIS\_2 \(\rightarrow\) FFFS\_2 (0.74) were identified.
In both instances, ``High'' FFFS was more likely when BIS was ``High'' (0.64 for \_1, 0.61 for \_2) compared to ``Low'' (0.33 for \_1, 0.37 for \_2).
The arc between nlelg \(\rightarrow\) BIS had a strength of 0.55 for nlelg\_1 \(\rightarrow\) BIS\_1, and 0.37 nlelg\_2 and BIS\_2.
``Low'' negative life events increased the probability of BIS being in the ``High'' state from 0.33 to 0.55 for nlelg\_1 \(\rightarrow\) BIS\_1, and 0.38 to 0.58 for nlelg\_2 \(\rightarrow\) BIS\_2.

\newpage

\begin{figure}
\centering
\includegraphics{apa_style_files/figure-latex/fig6-1.pdf}
\caption{\label{fig:fig6}The full Bayesian network structure was plotted using the \emph{strength.plot} function in bnlearn. The strength of each arc is shown graphically by the style of the arc. Thin, dashed arcs indicate the weakest arcs (arc strength below 0.50), whereas thick solid arcs indicate the strongest arcs (arc strength of 1). White nodes in the network indicate the explanatory variables, blue nodes indicate T1\_1 and T2\_1 variables, and red nodes indicated T2\_2 and T3\_2 variables. The injured\_X nodes have been coloured gold as they are the main nodes of interest within the network.}
\end{figure}

@@

The first conditional probability query (CPQ) investigated the variables in the Markov blanket for injured\_1 (Fig \ref{fig:fig7}), which contained hours spent training per week (hours), negative life events (nlelg\_1), muscle stiffness (stiffness\_1), current competitive level (clevel) and balance (balance\_1).
The arc between nlelg\_1 and injured\_1 was fixed in the network, so has the maximum strength of 1.

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{figures_doc/Fig7}

}

\caption{Markov blanket of injured\_1. Arc strengths are included as arc labels.}\label{fig:fig7}
\end{figure}

The CPQ for injured\_1 in the ``injured'' state for all variables that were linked to injured\_1 is shown in Table \ref{tab:table5}.
The probability of injured\_1 = ``injured'' rose from 0.17 to 0.31 when stiffness was ``High'' compared to ``Low''.
Negative life events had a negligible effect when moving from the ``Low'' to ``High'' state.

@@

The study investigated the multifaceted interactions of stress-related variables and injury occurrence using the stress-injury frameworks presented by Williams and Andersen (1998) and Appaneal and Perna (2014) and a prospective, repeated measures design applied to a large cohort of athletes. Relationships between stress-related psychosocial and physiological factors and injury were investigated using two BN structures; the first was a two-time Bayesian Network that investigated the relationships between variables across time points in the study, and the second network used differential equations to model the changes in variables between time points.
The latter facilitated the development of new insights into the interactions of stress-related factors with injury occurrence, exploring changes in both psychological and physiological factors that may occur preceding an injury.
